---
title: 'Stats 250: Case Study 1 Resources'
author: "Instructional Team"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\newpage

## Learning Objectives

1. Run a one-sample z-test for a proportion
2. Run a one-sample t-test for a mean


***


## Coding Examples

### Load Packages

Before we get started, be sure to load the following packages. 

```{r load_packages, message = FALSE, warning = FALSE}
library(ggplot2)
library(stats250package)
```

Note: the `stats250package` is a package created by the Stats 250 instructional team! It contains some helpful functions that we will use throughout the semester. 


***


### One Sample z-test for a Proportion

The city of Ann Arbor, which residents affectionately call "Tree Town", regularly assesses the health of public trees through a project call The Urban Forest. The data set below contains a random sample of 200 trees found in parks around Ann Arbor. Run the code chunks below to load and preview the data. 

```{r read_a2trees}
load("a2trees_sample.rda")
```

```{r preview_a2trees}
head(a2trees)
```

The variable `Health` categorizes trees as Very Good, Good, Neutral, Bad, or Very Bad. 

After a recent initiative to improve the health of trees, the city would like to assess if more than 30% of trees found in local parks are listed as being in "Good" or "Very Good" health. 

We can use the `table()` function to find how many of the 200 trees are listed as being in "Good" or "Very Good" health.

```{r frequency_table_example}
table(a2trees$Health)
```

We see that 16 + 57 = 73 trees fall into these categories.  

To test the idea statistically, we can use the `z.test()` function from our `stats250package`. The code would look like this:

```{r z_test_example}
z.test(x = 73, n = 200, pi = 0.30, alternative = "greater")
```

The arguments are as follows:

- `x`: the number of "successes"
- `n`: the sample size
- `pi`: the hypothesized proportion (from the null hypothesis)
- `alternative`: the direction of the alternative hypothesis ("greater", "less", or "two.sided")

The output provides us with:

- a summary of the data and the null probability
- the z-test statistic and corresponding p-value
- the alternative hypothesis that was tested
- a confidence interval (this can be ignored)
- the sample estimate (or pi_hat)

With a p-value of 0.02243, there is *strong* evidence against the null and in favor of the alternative. Based on the sample, there is *strong* evidence to suggest that over 30% of all trees in local Ann Arbor parks are listed as being in "Good" or "Very Good" health. 


***


### One Sample t-test for a Mean

The National Health and Nutrition Examination Survey (NHANES) is a program of studies designed to assess the health and nutritional status of adults and children in the United States. A random sample, representative of US adults aged 18+, is provided in the data set below. 

```{r read_nhanes}
nhanes <- read.csv("nhanes_sample.csv")
```

```{r preview_nhanes}
head(nhanes)
```

The variables include:

- `height`: height (in cm)
- `weight`: weight (in kg)
- `BMI`: body mass index (in kg/m^2)
- `arm`: arm circumference (in cm)
- `chol`: cholesterol level (in mg)
- `iron`: iron level (in mg)
- `hgb`: hemoglobin (in g/deciliter)
- `caffeine`: caffeine amounts (in mg)
- `rbcc`: red blood cell count (in millions of cells/microliter)
- `wbcc`: white blood cell count (in millions of cells/microliter)

In 2010, the average caffeine amount for US adults was reported to be 148 mg. We will use the provided sample to assess if the mean caffeine amount has changed since 2010.

To test this hypothesis, we can use the `t.test()` function in R. The code would look like this:

```{r t_test_example}
t.test(nhanes$caffeine, mu = 148, alternative = "two.sided")
```

The arguments are as follows:

- `variable$dataset`: the variable to be tested
- `mu`: the hypothesized mean (from the null hypothesis)
- `alternative`: the direction of the alternative hypothesis ("greater", "less", or "two.sided")

The output provides us with:

- the t-test statistic, degrees of freedom, and corresponding p-value
- the alternative hypothesis that was tested
- a confidence interval (this can be ignored)
- the sample estimate (or mu_hat)

With a p-value of 0.1421, there is *not enough* evidence against the null and in favor of the alternative. Based on the sample, there is *not enough* evidence to suggest that the average caffeine amount for US adults has changed (from 148 mg) since 2010.

These are the two new functions you will need to complete the case study. Now is a great time to ask questions! 


***


## Function Review

In order to complete the case study, you will need some functions from Labs 1 and 2. Instead of navigating back and forth, we have included some coding examples below. These examples will use the `penguins` data set. 

```{r read_penguins}
penguins <- read.csv("penguins.csv")
```

### Bar Chart

Create a bar chart to visualize a categorical variable. 

```{r bar_chart_example}
ggplot(data = penguins, aes(x = species)) + 
  
  geom_bar(color = "black", fill = "white") +
  
  labs(title = "Figure 1: Bar Chart of Penguin Species",
       x = "Species",
       y = "Frequency")
```

***

### Frequency Tables

Create a frequency table to summarize the counts of a categorical variable.

```{r table_example}
table(penguins$species)
```

***

### Histogram

Create a histogram to visualize the distribution of a quantitative variable. 

```{r histogram_example}
ggplot(data = penguins, aes(x = body_mass_g)) + 
  
  geom_histogram(bins = 25, 
                 color = "black", 
                 fill = "white") +
  
  labs(title = "Figure 2: Histogram of Penguin Body Mass",
       x = "Body Mass (in grams)",
       y = "Frequency")
```

***

### Numerical Summaries

Compute various numerical summaries for a quantitative variable, such as the mean, median, standard deviation, etc. 

```{r mean_example}
mean(penguins$body_mass_g)
```

```{r median_example}
median(penguins$body_mass_g)
```

```{r standard_deviation_example}
sd(penguins$body_mass_g)
```

```{r quantile_example}
quantile(penguins$body_mass_g)
```

```{r summary_example}
summary(penguins$body_mass_g)
```

Click on the file titled `casestudy01_report.Rmd` to navigate to the case study assignment!



